Smart Representations: {R}ationality, Regret, Evolution \& the Environment

### 2015-11-10 paper/paper.tex
Paolo references inserted

\usepackage[]{helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
% \bibliography{helpers/MyRefLocal}
\bibliography{choicePrince}
\usepackage{helpers/myenvironments}
\usepackage{helpers/mycommands}

@@

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modeling. However, most applications
look at evolutionary selection of behavior in a single \emph{stage game}. There are two common
simplifications in this approach. Firstly, the stage game is usually considered fixed, closed
and immutable. Secondly, what evolves are behavioral strategies for the stage game. It has
consequently been argued that, instead of focusing on a fixed, immutable stage game and the
behavioral selection for that stage game alone, we should rather study the evolutionary
competition of \emph{choice mechanisms} under selective pressure in a rich and variable
environment \citep[e.g.][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}.
But only little attention has been paid to modeling the evolutionary competition between choice
mechanisms in a game-theoretic framework \citep[but
see][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

If agents deal with a rich and varied environment, they face many different choice
situations. A crucial part of any choice mechanism is a \emph{subjective representation} of the
choice situation, e.g., the agent's manner of representing their preferences and her way of
forming decision relevant beliefs in the case of uncertainty. Since almost any reasonable
learning or decision-making process will make use of some (subjective, agent-internal (possibly
unconscious)) representation of the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central and has been
subject of recent interest (\textcolor{red}{CITE literature on evolution of
  preferences}). Na\"ive intuition might convince us that preferences that veridically
represent the actual, true fitness measure that evolutionary selection is based upon must
surely be the best. We demonstrate that, if we take into account how agents represent
uncertainty, this na\"ive conclusion is premature. Under certain reasonable epistemic
conditions, non-veridical preference representations can be evolutionarily beneficial.

@@

The standard textbook definition of \textit{rationality} in economics and decision theory
traces back to the seminal work of \citet{Neumannvon-NeumannMorgenstern1944:Theory-of-Games}
and \citet{Savage1954:The-Foundations}. It says that choice is rational only if it maximizes
(subjective) expected utility.

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. Prospect theory is, perhaps, the most prominent
example. \citet{KahnemannTversky1979:Prospect-Theory} looked at experimental data from human
choices in suitably defined tasks and demonstrated that what appears to be violations of
rationality norms can be explained on the assumption that subjects' beliefs and preferences
deviate systematically from the objectively given parameters in the presented choice task
(e.g., by taking into account reference points and diminishing returns for both gains and
losses).  Similarly, tinkering with the notion of subjective utility, in particular by
including social preferences such as fairness, reciprocity or similar, allows us to describe as
rational empirically observed behavior in situations of social decision making, such as the
prisoners' dilemma or public goods games, that might otherwise appear irrational (e.g. Fehr and
Schmidt, Charness and Rabin, etc.).\todo{@Paolo: please include proper references}

It could be objected that, without additional constraints, the notion of rationality is likely
to collapse, as it seems possible to deem rational everything that is observed, given the
freedom to adjust beliefs and preferences at will. But there are many ways in which ascriptions
of beliefs and preferences are constrained by normative considerations of rationality. To begin
with, if the DM's uncertainty is about objective chance, it can be argued that subjective
beliefs should conform (at least roughly) to objective chance. When the DM's uncertainty cannot
be anchored in an objective notion of chance, then temporal consistency in his decision making
should at least constrain his subjective beliefs, else he fall prey to exploitation by Dutch
books. Similarly, in certain situations it may make sense to postulate on normative grounds
that subjective preferences should be oriented towards tracking objective utility, e.g.,
fitness. For instance, in their classic work \myalert{(Alchian, Friedman)} argue that profit
maximization is a reasonable assumption for characterizing outcomes in a competitive market
because only firms behaving in a manner that is consistent with profit maximization will
survive in the long run.\todo{@Paolo: please add references}

An alternative to descriptive vs.~normative views on rationality is \emph{adaptationism}
\citep[e.g.][]{HagenChater2012:Decision-Making}. \todo{more references} Adaptationism aims to
explain choice behavior by appeal to evolutionary considerations. The key idea is that decision
makers (human and other life forms) have evolved choice mechanisms which were fitness enhancing
in the relevant domains of choice making, at the relevant time of evolution. When it comes to
preferences, this is clearly related to Alchian and Friedman's argument that firms
\emph{should} optimize that which makes them ``live,'' so to speak. For biological agents, the
argument would be that subjective preferences \emph{should} track whatever determines objective
fitness.

We would like to put this latter idea to the test. We present a formal argument intended to
support the conclusion that subjective preferences that differ from objective fitness can be
adaptive. We argue that it is important to take into account the potential mutual dependence of
subjective preferences and subjective beliefs in adaptationist approaches to decision
making. In other words, we focus on the evolutionary competition between \emph{subjective
  representation schemes}: general ways of forming subjective preferences and types of
subjective beliefs. We show that subjective, non-veridical preference representations based on
a notion of regret \citep[e.g.][]{Savage1951:The-theory-of-s,LoomesSugden1982:Regret-Theory:-}
can be adaptive and even outperform objective, veridical representations of the actual, true
fitness, when agents, at least on occasion, hold \emph{imprecise probabilistic beliefs}
(\textcolor{red}{CITATIONS}) and make decisions based on a security choice rule (here: the
maximin rule applied to expected utility). The next section introduces all of these notions
(precise/imprecise beliefs, veridical/regret representations, \dots) in detail.

@@

A strategic game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i$ a payoff function for each player $i \in N$. For the
most part, this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(\myact{k}, \myact{l}) = \pi_2(\myact{l}, \myact{k})$ for all $k$ and $j$. An example is
given in Figure~\ref{coordgame1}, which shows a coordination game. The numbers give the payoff
for the player choosing a row act against an opponent choosing a column act. Since the game is
symmetric, it suffices to give just these numbers.

@@

Rational choice from maximization of expected utilities, as given in
Definition~\ref{def:rationality}, is \emph{not} a choice rule in the present sense. Although
the notion applies to game situations, of course, it requires an additional component, namely a
belief of the DM, before it can say which acts the DM should choose. Let
$A_{-i} = \bigtimes A_j$ for all $j \neq i$ be the set of all \emph{profiles} of acts for all
players except $i$. If player $i$ has a belief $\mu_i \in \Delta(A_{-i})$ about the behavior of
his opponents, he chooses rationally only if his choice $a^*_i$ maximizes expected utility
$\EU(a) = \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. So this
requires a probabilistic belief about the opponents' action choices. In the example from
Figure~\ref{coordgame1}, a rational player would choose act $I$ exactly when he believes that
his opponent will choose act $I$ with a probability at least $\nicefrac{2}{3}$.

Perhaps the most well-known choice rule that does not require explicitly stating the DM's
beliefs is the \emph{maximin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act; players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

Another well-known choice rule is \emph{Laplace's rule} which prescribes to choose $\act^*$ if
$\act^* \in \argmax_{a \in A} \sum_{a'} \pi(a, a')$. Laplace's rule is equivalent to rationally
choosing under a \emph{flat belief} that assigns equal probability to all of the opponent's
choices. In the coordination game of Figure~\ref{coordgame1}, Laplace's rule selects act $II$.

The final choice rule that is relevant for our purposes here is \emph{regret
  minimization}. Regret theory has been introduced in decision theory by
\citet{Savage1951:The-theory-of-s} and Niherens [], and later developed by Bell [1982], Loomes
and Sugden [1982], and Fishburn [1982] independently.\todo{fix references}\ Lately,
\citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
solutions to game theoretical puzzles (like the Traveller's dilemma) in a way that is closer to
everyday intuition and empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:

@@

Formally, given a class $ \mathcal{G} $ of strategic games,
$\gamma_i: \mathcal{G} \rightarrow \Theta_i \times E_i$ is a function that assigns a subjective
preference $\theta_i \in \Theta_i$ and an epistemic state $e_i \in E_i$ to any game
$G \in \mathcal{G}$ for some agent $i$. A subjective preference $\theta_i$ for game $G$ and
agent $i$ is just a payoff function for $i$ suitable to the acts and co-players in
$G$. Crucially, this subjective payoff function can be different from the actual payoff in $G$,
which we interpret as the objective fitness. An epistemic state $e_i$ for game $G$ and agent
$i$ is some specification of $i$'s uncertainty about the co-players' behavior. Finally, an
action selection function for agent $i$ is a function
$\kappa_i: \Theta_i \times E_i \rightarrow \pow{A_i} $, that associates any subjective
representation of the game $ (\theta_i,e_i) \in \Theta_i \times E_i$ with a set of prescribed
actions for $i$ in $G$ under $i$'s subjective representation of $G$. We drop the index $i$ for
readability in the following.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In the case of the bag on my desk, a strict Bayesian, in the absence of any further information, would probably rely on the \textit{principle of insufficient reason} and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach [e.g., Walley91-96, Levi74, GardenforsANDSahlin82]. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant in a game theoretical context. Indeed, in a recent paper [Battigalliet al.15] Battigalli et al. write

@@

Hence, we consider two possible epistemic states in the face of uncertainty. The belief type $\Delta(A_{-i})$ represents a situation of non-Bayesian radical uncertainty, where the players are not able to narrow down the set of probability distributions to a single one and they consider all of them as possible. The other belief type $\overline{\mu}$ conceptualizes the uncertainty about the population by using the principle of insufficient reason and ascribes equal probability to all the possible alternatives. As we have seen, both options are reasonable and justifiable ways to deal with situations of uncertainty [REFERENCES]. \\

In decision theory, the imprecise probability model is in line with most of the representation results of decision making under uncertainty [GilboaSchmeidler89,KMM06], and seems justified by empirical observations too.

A prominent example is Ellsberg's paradox. A bag contains 90 marbles, 30 are red and the remaining 60 are either black or yellow. A marble is to be drawn and DM can win 100€ if she guesses the color of the drawn marble. The DM is offered two different bets. In the first one she can choose to bet either on red (R) or on yellow (Y), while in the second she can bet either on red or black (RB) or on yellow or black (YB), as shown in Table \ref{Ellsberg}.

@@

[[\noindent Definition \ref{def:expected fitness} gives the expected evolutionary fitness of player $(\tau, e)$ against player $(\tau', e')$ in game $G$.
[OF COURSE we will postpone the definition of evolutionary fitness, it is only momentarily here!]
\begin{definition} \label{def:expected fitness}

Given a game $G \in \mathcal{G}$, a player $(\tau, e)$, and a player $(\tau', e')$, we define $\pi^G((\tau, e),(\tau', e')):= \pi^G(\kappa(\tau(G), e),\kappa(\tau'(G), e'))$.

\end{definition}
[OF COURSE we will postpone the definition of evolutionary fitness, it is only momentarily here!]]]

It is then worth stressing here that $\kappa(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the actual payoff
maximinimizer, $\kappa(\tau^{reg}, \Delta(A_{-i}))$ corresponds to the regret minimizer,
$\kappa(\tau^{\pi}, \overline{\mu})$ is the expected actual payoff maximizer (with flat beliefs), and
$\kappa(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer (with flat beliefs).
Moreover, different subjective preferences and belief types can give rise to identical action predictions,
either in a restricted class of games, or in general. This is also the case here.

@@

\begin{example}
  Consider the game in Figure \ref{coordgame1} again. When a player characterized
  by\footnote{To ease notation in the following we will refer to a player directly as a pair
    $(player \text{ } type \text{ } \tau,\text{ } epistemic\text{ } type \text{ } e)$.} player
  type $ \tau^{reg} $ and epistemic type $ \Delta(A_{-i}) $ faces the game in Figure
  \ref{coordgame1}, her player type gives rise to the representation $\tau^{reg}(G^C)$ in
  Figure \ref{coordgame1reg}. Then, maximinimizing expected utility over the probability
  simplex $ \Delta(A_{-i}) $ given the game in Figure \ref{coordgame1reg} corresponds to
  playing the standard maxmin strategy [REF] of that game.\footnote{Notice that this is the
    case for any game: given a game $G$, maximinimizing expected utility over the probability
    simplex $ \Delta(A_{-i}) $ amounts to playing the standard maxmin strategy of $G$.} This,
  in turn, corresponds to the regret minimizing action of the original game $G^C$ of Figure
  \ref{coordgame1}, namely action $II$, as we have seen in Section \ref{sec:regreTheory}. A
  player $(\tau^{reg},\overline{\mu})$ instead has subjective representation $\tau^{reg}(G^C)$
  of Figure \ref{coordgame1reg} and a flat probability measure
  $\overline{\mu}=(\frac{1}{2}I,\frac{1}{2}II)$ over the co-player's actions. Consequently,
  such a player would play action $II$, i.e., the action that maximinimizes expected utility
  given the flat probability measure $\overline{\mu}$.\footnote{As we already noticed,
    maximinimizing expected utility given a probability measure $\mu$ is equivalent to
    maximizing expected utility given $\mu$.}  Similarly, a player
  $(\tau^{\pi},\Delta(A_{-i}))$ will have the subjective representation coinciding with the
  objective fitness game, i.e., $ \tau^{\pi}(G^C)= G^C $. Then, both actions $I$ and $II$
  maximinimize expected utility over the probability simplex $ \Delta(A_{-i}) $. Hence, a
  player $(\tau^{\pi},\Delta(A_{-i}))$ will be indifferent between $I$ and $II$, and will
  randomize with equal probability $(\frac{1}{2}I,\frac{1}{2}II)$. Finally, a player
  $(\tau^{\pi},\overline{\mu})$ maximinimizes the objective fitness game $G^C$ of Figure
  \ref{coordgame1} with flat probability $\overline{\mu}$ over the co-player's actions, i.e.,
  maximizes objective utility given $\overline{\mu}$. In game $G^C$ this yields action $II$.

@@

Evolutionary game theory has established itself as a useful tool for testing consequences of
sets of assumptions about complex evolutionary processes. However, the standard approach in
evolutionary game theory is not enough for our purposes, because it (i) fixes a single
\emph{stage game} (e.g., the prisoner's dilemma) and (ii) assumes that the object of
evolutionary selection is behavior for that very game. Following a similar line of argument
\citep[e.g.][]{FawcettHamblin2013:Exposing-the-be}, we therefore turn towards the evolutionary
competition of \emph{choice mechanisms}, i.e., general ways of selecting concrete behavior
across a range of different situations. As a consequence, the expected fitness of a choice
mechanism will not be determined by only a single (elementary) stage game, but it will be a
complex function of the statistical properties of the environment: we have to take into account
how frequently particular concrete situations of choice making are.

For concreteness, let us fix a class of strategic games $\mathcal{G}$ and a probability measure
$P \in \Delta(\mathcal{G})$ that gives the occurrence probability $P(G)$ of each game
$G \in \mathcal{G}$. Together, a class of games and a probability measure over it can be
thought of as a model of the environment in which different choice mechanisms compete. Let $C$
the a set of choice mechanisms that we care about. Then each choice mechanism $c \in C$ is a
associated with an \emph{action selection function} $S_c$ that maps any $G \in \mathcal{G}$
onto some action $S_c(G)$ of game $G$.\footnote{In all subsequent formal and simulation results
  we assume that, whenever a choice mechanisms is selects indifferently a set of actions, the
  agent will realize each with the same (flat) probability. For ease of parlor, we often do as
  if there was a tie-break rule.} Different choice mechanisms can select the same act in some
games. More importantly even, different choice mechanisms can have exactly the same action
selection functions on a particular class of games $\mathcal{G}$. This is because we think of
the object of evolutionary selection, not as the realized behavior but as the way behavior is
specified. (We will see below that this can make an important difference, e.g., in the light of
mutation.)

>>>>

\usepackage{etex}
\usepackage[nobiblatex]{helpers/mypackages}
\usepackage[backend=bibtex,natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{choicePrince,biblio}
\usepackage{helpers/myenvironments}
\usepackage{helpers/mycommands}

@@

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modelling. However, most applications
look at evolutionary selection of behavior in a single \emph{stage game}. There are two common
simplifications in this approach. Firstly, the stage game is usually considered fixed, closed
and immutable. Secondly, what evolves are behavioral strategies for the stage game. It has
consequently been argued that, instead of focusing on a fixed, immutable stage game and the
behavioral selection for that stage game alone, we should rather study the evolutionary
competition of \emph{choice mechanisms} under selective pressure in a rich and variable
environment \citep[e.g.][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. 
But only little attention has been paid to modeling the evolutionary competition between choice
mechanisms in a game-theoretical framework \citep[but
see][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

If agents deal with a rich and varied environment, they face many different choice
situations. A crucial part of any choice mechanism is a \emph{subjective representation} of the
choice situation, e.g., the agent's manner of representing their preferences and her way of
forming decision relevant beliefs in the case of uncertainty. Since almost any reasonable
learning or decision-making process will make use of some (subjective, agent-internal (possibly
unconscious)) representation of the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central and has been
subject of recent interest \citep[e.g.][]{algweib13,DekElyYlan07,Sam01,RobSam11}. Na\"ive intuition might convince us that preferences that veridically
represent the actual, true fitness measure that evolutionary selection is based upon must
surely be the best. We demonstrate that, if we take into account how agents represent
uncertainty, this na\"ive conclusion is premature. Under certain reasonable epistemic
conditions, non-veridical preference representations can be evolutionarily beneficial.

@@

The standard textbook definition of \textit{rationality} in economics and decision theory
traces back to the seminal work of \citet{deFinetti37}, \citet{Neumannvon-NeumannMorgenstern1944:Theory-of-Games}
and \citet{Savage1954:The-Foundations}. It says that choice is rational only if it maximizes
(subjective) expected utility.

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. Prospect theory is, perhaps, the most prominent
example. \citet{KahnemannTversky1979:Prospect-Theory} looked at experimental data from human
choices in suitably defined tasks and demonstrated that what appears to be violations of
rationality norms can be explained on the assumption that subjects' beliefs and preferences
deviate systematically from the objectively given parameters in the presented choice task
(e.g., by taking into account reference points and diminishing returns for both gains and
losses).  Similarly, tinkering with the notion of subjective utility, in particular by
including social preferences such as fairness, reciprocity or similar, allows us to describe as
rational empirically observed behavior in situations of social decision making, such as the
prisoners' dilemma or public goods games, that might otherwise appear irrational \citep[e.g.][]{fehrschmidt99,charrab02}.

It could be objected that, without additional constraints, the notion of rationality is likely
to collapse, as it seems possible to deem rational almost everything that is observed, given the
freedom to adjust beliefs and preferences at will. But there are many ways in which ascriptions
of beliefs and preferences are constrained by normative considerations of rationality. To begin
with, if the DM's uncertainty is about objective chance, it can be argued that subjective
beliefs should conform (at least roughly) to objective chance. When the DM's uncertainty cannot
be anchored in an objective notion of chance, then temporal consistency in his decision making
should at least constrain his subjective beliefs, else he fall prey to exploitation by Dutch
books. Similarly, in certain situations it may make sense to postulate on normative grounds
that subjective preferences should be oriented towards tracking objective utility, e.g.,
fitness. For instance, in their classic work \myalert{(Alchian, Friedman)} argue that profit
maximization is a reasonable assumption for characterizing outcomes in a competitive market
because only firms behaving in a manner that is consistent with profit maximization will
survive in the long run.\todo{@Paolo: please add references}

An alternative to descriptive vs.~normative views on rationality is \emph{adaptationism}
\citep[e.g.][]{HagenChater2012:Decision-Making}. \todo{more references} Adaptationism aims to
explain choice behavior by appeal to evolutionary considerations. The key idea is that decision
makers (human and other life forms) have evolved choice mechanisms which were fitness enhancing
in the relevant domains of choice making, at the relevant time of evolution. When it comes to
preferences, this is clearly related to Alchian and Friedman's argument that firms
\emph{should} optimize that which makes them ``live,'' so to speak. For biological agents, the
argument would be that subjective preferences \emph{should} stay as close as possible to whatever determines objective
fitness.

We would like to put this latter idea to the test. We present a formal argument intended to
support the conclusion that subjective preferences that differ from objective fitness can be
adaptive. We argue that it is important to take into account the potential mutual dependence of
subjective preferences and subjective beliefs in adaptationist approaches to decision
making. In other words, we focus on the evolutionary competition between \emph{subjective
  representation schemes}: general ways of forming subjective preferences and
subjective beliefs. We show that subjective, non-veridical preference representations based on
a notion of regret \citep[e.g.][]{Savage1951:The-theory-of-s,}
can be adaptive and even outperform objective, veridical representations of the actual, true
fitness, when agents, at least on occasion, hold \emph{imprecise probabilistic beliefs}
(\textcolor{red}{CITATIONS}) and make decisions based on a security choice rule (here: the
maximin rule applied to expected utility). The next section introduces all of these notions
(precise/imprecise beliefs, veridical/regret representations, \dots) in detail.

@@

A simultaneous-move game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i$ a payoff function for each player $i \in N$. For the
most part, this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(a_k, a_j) = \pi_2(a_j, a_k)$ for all $a_j,a_k \in A_1$. An example is
given in Figure~\ref{coordgame1}, which shows a coordination game. The numbers give the payoff
for the player choosing a row act against an opponent choosing a column act. Since the game is
symmetric, it suffices to give just these numbers.

@@

Rational choice from maximization of expected utilities, as given in
Definition~\ref{def:rationality}, is \emph{not} a choice rule in the present sense. Although
the notion applies to game situations, it requires an additional component, namely a
belief of the DM, before it can say which acts the DM should choose. Let
$A_{-i} = \bigtimes A_j$ for all $j \neq i$ be the set of all \emph{profiles} of acts for all
players except $i$. If player $i$ has a belief $\mu_i \in \Delta(A_{-i})$ about the behavior of
his opponents, he chooses rationally only if his choice $a^*_i$ maximizes expected utility
$\EU(a_i) := \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. So this
requires a probabilistic belief about the opponents' action choices. In the example from
Figure~\ref{coordgame1}, a rational player would choose act $I$ exactly when he believes that
his opponent will choose act $I$ with a probability at least $\nicefrac{2}{3}$.

Perhaps the most well-known choice rule that does not require explicitly stating the DM's
beliefs is the \emph{maximin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act: players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

Another well-known choice rule is \emph{Laplace's rule} which prescribes to choose $\act^*$ if
$\act^* \in \argmax_{a \in A} \sum_{a'} \pi(a, a')$. Laplace's rule is equivalent to expected utility maximization under a \emph{flat belief} that assigns equal probability to all of the opponent's
choices. In the coordination game of Figure~\ref{coordgame1}, Laplace's rule selects act $II$.

The final choice rule that is relevant for our purposes here is \emph{regret
  minimization}. The notion of regret in decision theory dates back at least to the work by
\citet{Savage1951:The-theory-of-s}, and has later been developed by \citet{bell82}, \citet{fish82} and \citet{loosug82} independently. Lately,
\citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
solutions to game theoretical puzzles (like the Traveller's dilemma) in a way that is closer to
everyday intuition and empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:

@@

Formally, given a class $ \mathcal{G} $ of strategic games,
$\gamma_i: \mathcal{G} \rightarrow \Theta_i \times E_i$ is a function that assigns a subjective
preference $\theta_i \in \Theta_i$ and an epistemic state $e_i \in E_i$ to any game
$G \in \mathcal{G}$ for some agent $i$. A subjective preference $\theta_i$ for game $G$ and
agent $i$ is just a payoff function for $i$ suitable to the acts and co-players in
$G$. Crucially, this subjective payoff function can be different from the actual payoff function $\pi^G$ of $G$,
which we interpret as the objective evolutionary fitness. An epistemic state $e_i$ for game $G$ and agent
$i$ is some specification of $i$'s uncertainty about the co-players' behavior. Finally, an
action selection function for agent $i$ is a function
$\kappa_i: \Theta_i \times E_i \rightarrow \pow{A_i} $, that associates any subjective
representation of the game $ (\theta_i,e_i) \in \Theta_i \times E_i$ with a set of prescribed
actions for $i$ in $G$ under $i$'s subjective representation of $G$. We drop the index $i$ for
readability in the following.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In the case of the bag on my desk, a strict Bayesian, in the absence of any further information, would probably rely on the \textit{principle of insufficient reason} and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach \citep[e.g.][]{levi74,gardsah82,walley96}. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant in a game theoretical context. Indeed, in a recent paper \citet{BattCerrMM15} write

@@

Hence, we consider two possible epistemic states in the face of uncertainty. The belief type $\Delta(A_{-i})$ represents a situation of non-Bayesian radical uncertainty, where the players are not able to narrow down the set of probability distributions to a single one and they consider all of them as possible. The other belief type $\overline{\mu}$ conceptualizes the uncertainty about the population by using the principle of insufficient reason and ascribes equal probability to all the possible alternatives. As we have seen, both options are reasonable and justifiable ways to deal with situations of uncertainty. \\

In decision theory, the imprecise probability model is in line with most of the representation results of decision making under uncertainty \citep[e.g.][]{gilsch89,KlibMarMuk05}, and seems justified by empirical observations too.

A prominent example is Ellsberg's paradox (\citet{ells61}). A bag contains 90 marbles, 30 are red and the remaining 60 are either black or yellow. A marble is to be drawn and DM can win 100€ if she guesses the color of the drawn marble. The DM is offered two different bets. In the first one she can choose to bet either on red (R) or on yellow (Y), while in the second she can bet either on red or black (RB) or on yellow or black (YB), as shown in Table \ref{Ellsberg}.

@@

It is then worth stressing that $\kappa(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the actual payoff
maximinimizer, $\kappa(\tau^{reg}, \Delta(A_{-i}))$ corresponds to the regret minimizer,
$\kappa(\tau^{\pi}, \overline{\mu})$ is the expected actual payoff maximizer (with flat beliefs), and
$\kappa(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer (with flat beliefs).
Moreover, different subjective preferences and belief types can give rise to identical action predictions,
either in a restricted class of games, or in general. This is also the case here.

@@

\begin{example}
  Consider the game in Figure \ref{coordgame1} again. When a player characterized
  by\footnote{To ease notation in the following we will refer to a player directly as a pair
    $(player \text{ } type \text{ } \tau,\text{ } epistemic\text{ } type \text{ } e)$.} player
  type $ \tau^{reg} $ and epistemic type $ \Delta(A_{-i}) $ faces the game in Figure
  \ref{coordgame1}, her player type gives rise to the representation $\tau^{reg}(G^C)$ in
  Figure \ref{coordgame1reg}. Then, maximinimizing expected utility over the probability
  simplex $ \Delta(A_{-i}) $ given the game in Figure \ref{coordgame1reg} corresponds to
  playing the standard maxmin strategy [REF] of that game.\footnote{Notice that this is the
    case for any game: given a game $G$, maximinimizing expected utility over the probability
    simplex $ \Delta(A_{-i}) $ amounts to playing the standard maxmin strategy of $G$.} This,
  in turn, corresponds to the regret minimizing action of the original game $G^C$ of Figure
  \ref{coordgame1}, namely action $II$, as we have seen in Section \ref{sec:choice-rules}. A
  player $(\tau^{reg},\overline{\mu})$ instead has subjective representation $\tau^{reg}(G^C)$
  of Figure \ref{coordgame1reg} and a flat probability measure
  $\overline{\mu}=(\frac{1}{2}I,\frac{1}{2}II)$ over the co-player's actions. Consequently,
  such a player would play action $II$, i.e., the action that maximinimizes expected utility
  given the flat probability measure $\overline{\mu}$.\footnote{As we already noticed,
    maximinimizing expected utility given a probability measure $\mu$ is equivalent to
    maximizing expected utility given $\mu$.}  Similarly, a player
  $(\tau^{\pi},\Delta(A_{-i}))$ will have the subjective representation coinciding with the
  objective fitness game, i.e., $ \tau^{\pi}(G^C)= G^C $. Then, both actions $I$ and $II$
  maximinimize expected utility over the probability simplex $ \Delta(A_{-i}) $. Hence, a
  player $(\tau^{\pi},\Delta(A_{-i}))$ will be indifferent between $I$ and $II$, and will
  randomize with equal probability $(\frac{1}{2}I,\frac{1}{2}II)$. Finally, a player
  $(\tau^{\pi},\overline{\mu})$ maximinimizes the objective fitness game $G^C$ of Figure
  \ref{coordgame1} with flat probability $\overline{\mu}$ over the co-player's actions, i.e.,
  maximizes objective utility given $\overline{\mu}$. In game $G^C$ this yields action $II$.

@@

The standard approach in
evolutionary game theory is not enough for our purposes, because it (i) fixes a unique and immutable
\emph{stage game} (e.g., the prisoner's dilemma) and (ii) assumes that the object of
evolutionary selection is behavior for that very game. Following a line of argument coming from theoretical biology and behavioral ecology
\citep[e.g.][]{FawcettHamblin2013:Exposing-the-be}, we therefore turn towards the evolutionary
competition of \emph{choice mechanisms}, i.e., general ways of selecting concrete behavior
across a range of different situations, defined by the two functions $\gamma$ and $\kappa$. With this approach, the expected fitness of a choice
mechanism will not be determined by a unique stage game, but it will be a
complex function of the statistical properties of the environment: we have to take into account
how frequently particular concrete situations of choice making are.

For concreteness, let us fix a class of strategic games $\mathcal{G}$ and a probability measure
$P \in \Delta(\mathcal{G})$ that gives the occurrence probability $P(G)$ of each game
$G \in \mathcal{G}$. Together, a class of games and a probability measure over it can be
thought of as a model of the environment in which different choice mechanisms compete. As we have seen in Definition \ref{def:maxminSEU}, given a game $G \in \mathcal{G}$, a choice mechanism prescribes the action(s) $\kappa(\gamma(G))$. Now imagine a huge, possibly infinite society of agents, each of which carries a choice mechanism, and a class $\mathcal{G}$ of symmetric two player games for simplicity.
Definition \ref{def:expected evolutionary payoff} gives the expected evolutionary payoff of choice mechanism $\kappa(\tau, e)$ against choice mechanism $\kappa(\tau', e')$ in game $G \in \mathcal{G}$.

\begin{definition} \label{def:expected evolutionary payoff}
